\chapter{Introduction}
\label{chap:introduction}

%\bit
%\item Stationary Points in general
%\item Finding extrema
%\item Other sorts of stationary points
%\item Minima in theoretical chemistry
%\item Saddle points in Chemistry (reaction rates)
%\item Why do we want to find higher order saddle points in chemistry.
%\item Energy storage, conversion and mobility (Hydrogen as an example)
%\item The importance of assessing reaction rates (Kinetics is a very important aspect for the above)
%\item General problems with getting reaction rates
%\item Annoying landscapes
%\item Beyond harmonicity
%\item The methods for finding stationary points presented here are inspired by chemistry but are applicable to any function.
%\item The thesis is heavily biased towards chemistry
%\eit

\section{Stationary Points}
The border between mathematics, physics and chemistry is often times obscure and the methods developed for one, may very well benefit the others.
This relationship is obvious for mathematical methods that get adopted for physics and chemistry but the reverse is also possible (for example the Davidson method for eigenvalue and eigenvector calculation~\cite{davidson-method-1975}).
Common to all the aforementioned, and other, fields, is some form of functional analysis where the study of stationary points\footnote{Points where the first derivative is zero.} is generally of great interest.
For systems of limited dimensionality, finding stationary points is no more complex than solving a root problem for the function's derivative.
However, once systems grow beyond a few dimensions, solving the root problem becomes infeasible.
Furthermore, the lack of a global analytical function makes it impossible to simply solve for the roots.
This is, generally, the case in simulations on the atomic scale, which is the main subject of this thesis.
While the methods discussed here are developed in that field, applying them to purely mathematical problems and within other fields is merely a question of interest and implementation, thus they are presented as general as possible in order to appeal to a wider audience.

Locating extrema using local information, i.e. the gradient, is a well known technique~\cite{cg-original-1952}, whereas the gradient is simply followed in the appropriate direction (positive for maxima and negative for minima) until it vanishes.\footnote{Global extrema are, conversely, extremely difficult to find using only local information and guaranteeing that a given extremum is a global one is nearly impossible.~\cite{Stillinger1999} See~\cite{Wang2010, Wales1999, Pickard2011, Oganov2008, Kirkpatrick1983, Johnston2003, Fischer2006} for such efforts.}
Finding other stationary points is less obvious and classifying them depends on the second order derivatives (or even higher in special cases\footnote{For example the Monkey Saddle \url{http://mathworld.wolfram.com/MonkeySaddle.html}})
Finding such points without explicitly evaluating higher order derivatives is an important subject within atomic simulations, since their calculation is very costly when using non-analytical models to describe the atomic interaction.

\section{Atomic Simulations}
The focus of this thesis is the application and improvement of theoretical methods to calculational reaction rate chemistry.
The underlying methods that allow this are numerous and their discovery span ages in time, from Newton's equations of motion~\cite{newton-latin} up to modern day methods for solving Sch\"odinger's equation~\cite{schrodinger-equation-1926} for quantum systems~\cite{hohenberg-kohn-1964, gpaw-review-2010, dacapo-1999}, from the discovery of calculus and vector math to modern methods for solving eigenvalue problems~\cite{eigenvalue-problems-2000}.

Simulations of atomic systems are an integral part of modern atomic scale research, whether it be to investigate specific electronic structure properties (e.g.~\cite{electronic-study}), performing atomic dynamics (e.g.~\cite{dynamics-review-2002}), investigating macroscopic properties (e.g.~\cite{macroscopic}) or simply finding stable structures (e.g.~\cite{summer-school-paper-2009}).
Simulations are commonly employed in the analysis of interesting material properties (see for example catalytic properties~\cite{catalytic}), the screening of candidate materials --- to eliminate the need to synthesize as many unsuccessful materials --- for various processes (see for example hydrogen storage~\cite{summer-school-paper-2009}) and further collaboration and comparison with experiments (see for example stability research of ammines~\cite{order-disorder-2010}).

Reaction chemistry, generally, happens on a timescale of microseconds which is, essentially, eternity when seen from the vantage point of the fast moving atomic vibrations, which happen on a timescale of femtoseconds.
Bridging this gap is an important research subject which remains open, despite noble efforts (some of which are described in \fref{chap:tst}) that have moved the field a long way by introducing statistical methods that are able to exploit the timescale difference to treat the events independently of each other~\cite{tst-eyring-1935, tst-polanyi-1935, tst-wigner-1938, htst-wert-1949, htst-vineyard-1957, tst-1990}.

\subsubsection{The Significance of Stationary Points in Atomic Simulations}
The potential energy of atomic systems can be evaluated as a function of the atomic coordinates~\cite{born-oppenheimer-1927, schrodinger-equation-1926, kohn-1999}.
The most important features of such mappings are generally considered to be the minima which represent stable configurations of the atoms.
The transition between minima is then considered to be a chemical reaction.
These can range, for example, from rotations of specific groups with minimal impact on the properties of the material (as seen in \fref{chap:borohydrides}) and motion of water molecules in liquid water, where the chemistry remains unchanged, to adsorption and the breaking of strong chemical bonds.
In understanding the transition in question, the path of least resistance (or minimum energy path) is an essential concept, since the highest point along it defines the potential energy that must be overcome and the path itself is a good indicator of the reaction pathway.
The top point is a non-extremal stationary point (in this case, a first order saddle point) and information about the frequency of a given transition (the reaction rate) can be inferred from from it and its surroundings, which is of central importance in chemistry~\cite{htst-wert-1949, htst-vineyard-1957}.
The surroundings of a first order saddle point are considered its immediate environment, but care must also be taken that any neighbouring higher order saddle points are sufficiently far away in potential energy and space.
This latter criteria for the surroundings is often disregarded, especially in systems for which the approximations in question generally perform well.
Part of the reason for this is the lack of a simple-to-use method for mapping the surroundings and neighbouring saddle points.
In this thesis an effort is made to bridge that gap (\fref{chap:erm}).

\section{Mobile Energy Solutions}
Being able to transport energy in a compact and safe way is a major concern of modern society.
Offering alternatives to fossil fuels is essential and thus, a highly active field of research is the design of novel materials for energy conversion, energy storage and catalysis.
If energy storage is taken as an example, given appropriate energy densities and stability, kinetics are essential.
Reaction rates are used, either directly or indirectly, to investigate all of these factors, naturally with particular emphasis on kinetics.

\subsubsection{Fossil Fuels}
Ancient lifeforms, chemically packed with the Sun's energy and compressed over millions of years by the weight of the Earth's crust, have served mankind cheap and readily available energy for millennia as fossil fuels~\cite{early-oil-use}.
When only considered as an energy carrier, fossil fuels are superb.
So superb, in fact, that any current alternative mobile storage medium is neither competitive with regards to energy density nor price.
This is, of course, not the whole story as the production of fossil fuels has been expected to reach its peak for some time~\cite{peak-oil-1956}.
Once this peak has been reached, the increasing demand of the ever expanding population, would quickly overwhelm the production capabilities and, more importantly, the Earth's reserves.
Fossil fuels will not be an option.
Furthermore, the ongoing climate change is an unacceptable side product of burning carbon rich fossils~\cite{climate-change-2007}.

Finding a viable alternative is one of the most important research subjects of the late 20th and early 21st centuries.

\subsubsection{Energy Trapping}
Fossil fuels serve both as energy sources and carriers.
If they are to be replaced alternatives must be found for both.
Sustainable ways to capture the Sun's energy tend to do so intermittently, e.g. solar panels can not be kept in direct sunlight during the night and at times the weather is still during which, windmills are of little use.
The problem is then to regulate the energy, to store excess when it is produced and to distribute it when none is produced.
This is where the fossil fuels serve(d) as energy carriers.

\subsubsection{Hydrogen as an Alternative}
Out of all the elements, hydrogen is the lightest and its chemical bonds hold a remarkable amount of energy.
Its abundance in on Earth and its non-harmful by-products (water and oxygen), make it a great candidate for a mobile energy store.

Directly storing hydrogen as a gas requires a very large volume or high pressure while storing it as a liquid requires cooling and/or pressurised tanks.
Neither is a very compelling solution for use in, for example, the average car.
Such solutions have, nevertheless, been deployed for example in buses where it is easy to fit huge tanks. %~\citemiss

Another way of storing hydrogen is in metal hydrides, where the hydrogen is either trapped in the metal's lattice or chemically bound within the metal.
These storage solutions tend to have more reasonable volumetric densities (at a reasonable cost of the gravimetric density) but often require excessive operating conditions for loading/unloading~\cite{borohydride-stability-2006}.

The reasons for the low volumetric capacity of the pure hydrogen solutions is simple, the tank is too big.
But understanding the underlying reasons for the high stability is an ongoing research topic.
An essential piece in that puzzle is understanding the kinetics of the material, how the hydrogen moves and how/why/when the material deforms.
Such knowledge can then be used to create better materials with more appropriate properties.

\subsubsection{Fuel Cells}
Once hydrogen has been stored and retrieved it is generally used in fuel cells to produce electricity without combustion. %\citemiss
Currently this process depends on rare and expensive elements, such as platinum, to catalyse the reaction.
This dependence on low-availability materials can very well become a bottleneck for the implementation of the so-called hydrogen economy.
Investigating catalytic activity and designing cheap, efficient, catalysts is highly dependant on understanding the kinetics and the catalytical pathways involved.

\section{Summary}
Most of the challenges the hydrogen economy faces are limited by kinetics in some way or other.
Catalysis and storage currently being the most important ones.

Near saddle points is encoded information about kinetics which is highly useful in designing and implementing the future.
Helping to understand the problems of the past and designing novel materials for the future.

%\item Why statistical methods
%\item Why path techniques (introduction to be expanded on in the TST section)
%\item It is now "possible" to do long timescale MD simlations but statistical methods are still more suited to find "all" the processes (chat with Elvar on MD)
%\item (Why these methods but not some other?)

%\section{Notation}
%\bit
%\item Nuclei positions: $\vR$
%\item Electron positions: $\vr$
%\item R-basis for formulas
%\item The PES: $E(\vR)$
%\item Gradient vs. Force
%\eit

\input{chapters/introduction/chapters}
