\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

\placeholder
\begin{comment}
At the end of my Ph.D. project, I would like to express my thanks many people.

First of all are my supervisors, Tejs Vegge and Hannes J\'onsson.
Tejs' ever positive attitude, always seeing the positive side in all situations served as an incredible inspiration to me both with regards to the project and life in general.
Every meeting of ours would leave me in a better place.
Hannes' involvement was essential to the project.
He is such a fountain of knowledge with regards to all aspects of my project.
I enjoyed our discussion of the sciences tremendously.

Then come my co-workers at DTU and Ris\o{} who are literally too many to appropriately thank each one.
The memers of Tejs Vegge's group who have come and gone.
My office buddies at DTU.

My trusty team of proofreaders and fellow group members warrant special thanks: Jakob Howalt, Steen Lysgaard, J\'on Steinar Gar\dh{}arsson M\'yrdal and Peter Jensen. Cake will be served.

Special thanks go to my lunch buddy and right-hand man Elvar \"Orn J\'onsson and his posse of Icelanders.

My friends, parents, grandparents and seemingly endless amount of siblings have a special place in my heart for tolerating long bouts of my stays abroad.

And finally my main buddy in chemistry (studies and that of the heart), Ester, I love you!
\end{comment}

%\begin{itemize}
% \item Professor Hannes J\'onsson, at the University of Iceland, for his guidance and inspiration.
% \item Associate Professor Yoshitada Morikawa, at the University of Osaka, for his guidance and kind manner while I was in a strange culture.
% \item Egill Sk\'ulason for collaboration, discussions and proof-reading.
% \item The members of the J\'onsson research group, at the University of Iceland, for enlightening discussions and support:
% \begin{itemize}
%  \item Sigr\'i\dh{}ur Gu\dh{}mundsd\'ottir, also for proof-reading
%  \item J\'on Steinar Gar\dh{}arsson M\'yrdal
%  \item Andreas Pedersen
%  \item Jean-Claude Berthet, Ph.D.
%  \item Peter Kl\"upfel, Ph.D.
% \end{itemize}
% \item The members of the J\'onsson research group, at the University of Iceland, for help with compilation issues, discussions and proof-reading:
% \begin{itemize}
%  \item Andri Arnaldsson, Ph.D.
%  \item Finnbogi \'Oskarsson
% \end{itemize}
% \item The members of the Yoshida research group, at the University of Osaka, for discussions, help and kind acceptance in a strange culture:
% \begin{itemize}
%  \item Susumu Yanagisawa, Ph.D.
%  \item Ikutaro Hamada, Ph.D.
%  \item Kunihiko Yamauchi, Ph.D.
% \end{itemize}
%% \item The best curry place in existence, Senba Curry, at the train station at Kita-Senri in Osaka, Japan.
% \item Friends and family for putting up with my studies for such a long time.
% \item Computer facilites: Bj\'olfur, J\"otunn and Itanium
% \item Funding:
% \begin{itemize}
%  \item Japanese Government (MONBUKAGAKUSHO:MEXT) Scholarship
%  \item Science Institute of the University of Iceland
% \end{itemize}
%\end{itemize}
