\section*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This thesis is submitted in candidacy for the Ph.D. degree from the Technical University of Denmark (DTU).
The work was been carried out between November 2008 and February 2012 at the Ris\o{} DTU campus, formerly the Ris\o{} National Laboratory for Sustainable Energy, DTU; the Center for Atomic-scale Materias Design (CAMD) and with external stays at the Science Institute of the University of Iceland.
The project was supervised by Senior Scientist Tejs Vegge, Head of Section at the Department of Energy Conversion and Storage at the Technical University of Denmark and co-supervised by Professor Hannes J\'onsson, Vice-Director of the Chemistry Division of the Science Institute of the University of Iceland.
The work was financially supported by DTU and CAMD, as well as the Nordic Center of Excellence on Hydrogen Storage Materials, the European Graduate School for Sustainable Energy Technologies - The Molecular Approach and Catalysis for Sustainable Energy (CASE).

\vspace{1em}

At the end of my Ph.D. project, I would like to express my thanks to the many people who made my project possible and my stay in Denmark an enjoyable one.

First of all are my supervisors, Tejs Vegge and Hannes J\'onsson.
Tejs' ever positive attitude, always seeing the positive side in all situations served as an incredible inspiration to me both with regards to the project and life in general.
Every meeting of ours would leave me in a better place.
Hannes' involvement was essential to the project.
He is such a fountain of knowledge with regards to all aspects of my project.
I enjoyed our discussion of the sciences tremendously.

Then come my co-workers at DTU and Ris\o{} who are literally too many to appropriately thank each one.
The members of Tejs Vegge's group who have come and gone.
My office buddies at DTU and those down the hall.
I thank you all for the interesting and entertaining discussions throughout my stay.
My trusty team of proofreaders and fellow group members warrant special thanks: Jakob, Steen, J\'on Steinar and Peter.
Cake will be served.

Special thanks go to my lunch buddy and right-hand man Elvar and his posse of Icelanders.
As well as to Da\dh{}i for introducing me to the cinematic pleasures of The Room.
D\'undri, Bobb\'y, Hulda, \'Oli, Habb\'y, Biggi, KM\&P, Little Billy Hanning and  Yury all did their best at keeping my sanity intact.
My friends, parents, grandmothers and seemingly endless amount of siblings have a special place in my heart for tolerating long bouts of my stays abroad.

And finally my aptly named buddy in chemistry and life, Ester, I love you!

%And finally my main buddy in chemistry (studies and that of the heart), Ester, I love you!

\vspace{1em}
\begin{flushright}
Kgs. Lyngby, Denmark, February 29th 2012,\\

\vspace{2em}

\rule{50mm}{0.1pt}\\
\textit{J\'on Bergmann Maronsson}
\end{flushright}
